---
title: "visualizations"
format: html
editor: visual
---

## Import Spatial Data

Import spatial data. Geopackage was preprocessed using Quantum GIS Version 3.34.5.

```{r import spatial data}

#| echo: false
#| warning: false
#| message: false
#| results: 'hide'

## All Spatial Data is stored in a geopackage called basic_data. Basic_data consists of several layers such as the street network, surface type or housing footprint. The layers were preprocessed and cliped to the extent of the city outline of zurich using Quantum GIS Version 3.34.5


## Show layers in basic_data.gpkg
st_layers("gis_files/basic_data.gpkg")


## Import street network from zurich, based on the swisstlm3d
streets <- read_sf("gis_files/basic_data.gpkg", "street_network_z") |> 
   select(objektart, geom) |> 
  mutate(
    objektart = as.factor(objektart),
    width = as.numeric(substr(objektart, start = 1, stop = 1)),
  ) |> 
  na.omit()


## Import city border of zurich
outline <- read_sf("gis_files/basic_data.gpkg", "city_outline")


## Import housing footprint of zurich
housing <- read_sf("gis_files/basic_data.gpkg", "housing_footprint") |> 
   select(objektart, geom) |>
  mutate(
    objektart = as.factor(objektart)
  )


## Import surface type of zurich
surface <- read_sf("gis_files/basic_data.gpkg", "surface_type") |> 
   select(art, geom) |> 
  mutate(
    art = as.factor(art)) |> 
 filter(art == "fliessendes Gewässer" | art == "stehendes Gewässer" | art == "Strasse, Weg" | art == "Verkehrsinsel") |> 
  na.omit()


## Import digital height model DHM25, a set of data representing the 3D form of the earth’s surface without vegetation and buildings
height <- terra::rast("gis_files/dhm25_zh.tif")


```

## 1. Visualize raw Data

Visualize raw GPS data of both messengers using a simple ggplot.

```{r visualize raw data via ggplot}

load("rda_files/all_routes.rda")
load("rda_files/raeubertochter_raw.rda")
load("rda_files/donner_raw.rda")


## Visualize the raw gps data from each messenger


### Raeubertochter
p_raw_raeubertochter <- ggplot() +
  geom_sf(data = outline, alpha = 0) +
  geom_sf(data = raeubertochter_raw, aes(color = "red"), size = 0.6) +
  coord_sf(datum = st_crs(2056)) +
  labs(x = "E",
       y = "N") +
    theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 90),
    axis.title.y = element_text(margin = margin(r = 20)),
    axis.title.x = element_text(margin = margin(t = 20))
  )

ggsave(plot = p_raw_raeubertochter, "figures/raeubertochter_raw.png", width = 150, height = 125, dpi = 300, units = "mm")


### Donner
p_raw_donner <- ggplot() +
  geom_sf(data = outline, alpha = 0) +
  geom_sf(data = donner_raw, aes(color = "red"), size = 0.6) +
  coord_sf(datum = st_crs(2056)) +
  labs(x = "E",
       y = "N") +
  theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 90),
    axis.title.y = element_text(margin = margin(r = 20)),
    axis.title.x = element_text(margin = margin(t = 20))
  ) 


ggsave(plot = p_raw_donner, "figures/donner_raw.png", width = 150, height = 125, dpi = 300, units = "mm")



```

Subsample of raw data in a smaller spatial context

```{r visualize raw data via tmap and spatial context}

## Next, we want to visualize a subsample of our raw data in a smaller spatial context
## visualize Subsample in spatial context


## fixing an extent for a subsample of raeubertochter
coords <- rbind(c(2681695, 1247822), c(2683188, 1247822), c(2683188, 1246616), c(2681695, 1246616), c(2681695, 1247822))


## create an extent polygon
polygon <- st_polygon(list(coords))


## convert to spatial object
extent_raeubertochter <- st_sfc(polygon, crs = 2056)


## clip spatial data to extent to reduce data points and speed up computing
housing_raeubertochter <- st_intersection(housing, extent_raeubertochter)

# housing_raeubertochter <- housing_raeubertochter |> 
#   group_by(objektart) |> 
#   summarise(geometry = st_union(geom)) |> 
#   ungroup()

streets_raeubertochter <- st_intersection(streets, extent_raeubertochter)
surface_raeubertochter <- st_intersection(surface, extent_raeubertochter)
raeubertochter_ext <- st_intersection(raeubertochter_raw, extent_raeubertochter)



## Reduce data set to every second gps fix to get a better overview
subset_raeubertochter_raw <- raeubertochter_ext[seq(1, nrow(raeubertochter_ext), by = 2), ]


## map including spatial context of zurich

tmap_mode("plot")
tm_raw_context <- tm_shape(streets_raeubertochter, bbox = extent_raeubertochter) +
  tm_lines(lwd = "width", scale = 2.5, legend.lwd.show = FALSE) +
tm_shape(housing_raeubertochter, bbox = extent_raeubertochter) +
  tm_polygons(col = "grey25", border.col = "white", lwd = 0.5, alpha = .9) +
tm_shape(subset_raeubertochter_raw, bbox = extent_raeubertochter) +
  tm_symbols(size = 0.6, col = "red",  alpha = 0.7, border.col = "red", legend.col.show = FALSE) +
  tm_layout(frame = FALSE)


## show map
print(tm_raw_context)


## save as a png
tmap_save(tm_raw_context, "figures/raeubertochter_raw_context.png", width = 250, height = 200, dpi = 300, units = "mm")

```

Visualize raw data via interactive map

```{r visualize raw data via interactive map}

## Browse trough raw_data

raw_data <- all_routes[seq(1, nrow(all_routes), by = 2), ]

tmap_mode("view")

tm_shape(raw_data) +
  tm_symbols(size = .25, col = "messenger",  alpha = 0.7, legend.col.show = FALSE, palette = c("red", "yellow"), border.col = "black", border.lwd = 0.1) +
  tm_layout(frame = FALSE) +
  tm_basemap(server = "CartoDB.PositronNoLabels", alpha = 0.75) +
  tm_view(symbol.size.fixed = F) +
  tm_facets(by = "messenger", sync = TRUE, ncol = 2)


######################################### LEAFLET #################################

raw_data |> 
  st_transform(crs = 4326) |>
leaflet() |>
    addCircleMarkers(opacity = 0.5,
                     color = ~pal(messenger),
                     radius = 2,
                     label = ~as.character(messenger),
                     group = ~messenger) |> 
    addProviderTiles(providers$CartoDB.PositronNoLabels, options = tileOptions(opacity = 0.75))  |> 
  addScaleBar(position = "bottomleft", scaleBarOptions(imperial = FALSE)) |> 
  addLegend(position = 'topright',
            pal = pal,
            values = ~messenger,
            title = "bike messengers")  |> 
  addLayersControl(~messenger, ) 
  


```

## 2. Visualize sampling interval

```{r visualize sampling intervals via ggplot}

## Quick overview how time differences distribute
p_sampling_interval <- ggplot(all_routes, aes(x = time_difference)) +
  geom_histogram(binwidth = .25, col = "white", fill = "grey25") +
  scale_y_log10(labels = ~ format(.x, scientific = FALSE)) +
  scale_x_log10(labels = ~ format(.x, scientific = FALSE)) +
  labs(x = "Time Difference in sec. (Log Scale)", y = "Count (Log Scale)") +
  facet_wrap(all_routes$messenger) +
  theme_minimal() +
  theme(
    axis.title.y = element_text(margin = margin(r = 20)),
    axis.title.x = element_text(margin = margin(t = 20)),
    plot.title = element_text(margin = margin(b = 10))
  )


# Both with similar but slightly different sampling regimes, raeubertochter with an intervall of ~1 second. It seems that static time is already removed in here dataset -> strava data! Most time differences between 1 and 10 seconds, some outliers at more than 1000 seconds, mut most of the points within 10s

# Donner with more variation between fixes, but still short intervals, might need another segmentation...

## show histogram
print(p_sampling_interval)

ggsave(plot = p_sampling_interval, "figures/sampling_intervall.png", width = 300, height = 150, dpi = 500, units = "mm")

```

## 3. Visualize Segmented Trajectories

Visualize segmented GPS data of both messengers using a simple ggplot.

```{r visualize segmented data via ggplot}

load("rda_files/all_routes_seg.rda")
load("rda_files/raeubertochter_seg.rda")
load("rda_files/donner_seg.rda")

## Visualize the segmented gps data from each messenger
### Raeubertochter

## create own color palette
my_palette <- colorRampPalette(brewer.pal(11, "Spectral"))


## visualiaze raeubertochter deliveries


### Raeubertochter
p_seg_raeubertochter <- ggplot() +
  geom_sf(data = outline, alpha = 0) +
  geom_sf(data = raeubertochter_seg, aes(color = segment_id), size = 0.6) +
  scale_color_manual(values = my_palette(22)) +
  coord_sf(datum = st_crs(2056)) +
  labs(x = "E",
       y = "N") +
    theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 90),
    axis.title.y = element_text(margin = margin(r = 20)),
    axis.title.x = element_text(margin = margin(t = 20))
  )

p_seg_raeubertochter
ggsave(plot = p_seg_raeubertochter, "figures/raeubertochter_seg.png", width = 150, height = 125, dpi = 500, units = "mm")


### donner
p_seg_donner <- ggplot() +
  geom_sf(data = outline, alpha = 0) +
  geom_sf(data = donner_seg, aes(color = segment_id), size = 0.6) +
  scale_color_manual(values = my_palette(36)) +
  coord_sf(datum = st_crs(2056)) +
  labs(x = "E",
       y = "N") +
    theme_minimal() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 90),
    axis.title.y = element_text(margin = margin(r = 20)),
    axis.title.x = element_text(margin = margin(t = 20))
  )


ggsave(plot = p_seg_donner, "figures/donner_seg.png", width = 150, height = 125, dpi = 500, units = "mm")




```

Subsample of segmented data in a smaller spatial context

```{r visualize segmented data via tmap and spatial context, fig.height=6, fig.width=8, fig.dpi=500}


### Next, we want to visualize a subsample of our segmented data in a smaller spatial context
## Visualize Subsample in spatial context


## Clip segmented data to a smaller extent to reduce computing time
raeubertochter_ext2 <- st_intersection(raeubertochter_seg, extent_raeubertochter)


## create smaller subsample consisting of every second point to get a better overview
subset_raeubertochter_seg <- raeubertochter_ext2[seq(1, nrow(raeubertochter_ext2), by = 2), ]


## transform segment_id to numeric and back to factor
subset_raeubertochter_seg$segment_id <- as.numeric(subset_raeubertochter_seg$segment_id)
subset_raeubertochter_seg$segment_id <- as.factor(subset_raeubertochter_seg$segment_id)
unique(subset_raeubertochter_seg$segment_id)

## Create own color palette
my_palette2 <- c("#9E0142", "#F46D43", "#FFFF66", "#66C2A5","#3288BD", "#5E4FA2")


tmap_mode("plot")


tm_seg_context <- tm_shape(streets_raeubertochter, bbox = extent_raeubertochter) +
  tm_lines(lwd = "width", scale = 2.5, legend.lwd.show = FALSE) +
tm_shape(housing_raeubertochter, bbox = extent_raeubertochter) +
  tm_polygons(col = "grey25", border.col = "white", lwd = 0.5, alpha = .9) +
tm_shape(subset_raeubertochter_seg, bbox = extent_raeubertochter) +
  tm_symbols(size = 0.6, col = "segment_id", palette = my_palette2, alpha = 0.7, border.col = "NA", legend.col.show = FALSE) +
  tm_layout(frame = FALSE) 


## Show map
print(tm_seg_context)


## save as png
tmap_save(tm_seg_context, "figures/raeubertochter_seg_context.png", width = 250, height = 200, dpi = 300, units = "mm")


```

Visualize raw data via interactive map

```{r visualize segmented data via interactive map}

seg_data <- all_routes_seg[seq(1, nrow(all_routes), by = 2), ]

tmap_mode("view")

tm_shape(raw_data) +
  tm_symbols(size = .25, col = "messenger",  alpha = 0.7, legend.col.show = FALSE, palette = "Spectral", border.col = "black", border.lwd = 0.1) +
  tm_layout(frame = FALSE) +
  tm_basemap(server = "CartoDB.PositronNoLabels", alpha = 0.75) +
  tm_view( symbol.size.fixed = F) +
  tm_facets(by = "messenger", sync = TRUE, ncol = 2)

```

## 4. Visualize Tunnel Merging

Visualize automated workflow for troubling tunnel problems

```{r visualize corrected segments}

load("rda_files/all_routes_seg.rda")

a_to_b <- all_routes_seg |> 
  group_by(segment_id) |> 
  slice(c(1, n())) |> 
  ungroup()

### we want to visualize the correction progress from the original segmentation to the corrected version including our tunnel problem

tmap_mode("plot")

## create a background for future maps
background <- tm_shape(streets_raeubertochter, bbox = extent_raeubertochter) +
  tm_lines(lwd = "width", scale = 2.5, legend.lwd.show = FALSE) +
tm_shape(housing_raeubertochter, bbox = extent_raeubertochter) +
  tm_polygons(col = "grey25", border.col = "white", lwd = 0.5, alpha = .9) +
   tm_layout(frame = FALSE) 


## First show segments
p1 <-  background + tm_shape(subset_raeubertochter_seg, bbox = extent_raeubertochter) +
  tm_symbols(size = 0.9, col = "segment_id", palette = my_palette2, alpha = 0.9, border.col = "NA", legend.col.show = FALSE) +
   tm_layout(frame = FALSE, bg.color = "transparent")

## save to png 
tmap_save(p1, "figures/raeubertochter_seg_context.png", width = 250, height = 200, dpi = 300, units = "mm")


## Then show start and ending points of trajectories from raeubertochter inside our extent
first_last_points_raeubertochter <- a_to_b |> 
  filter(messenger == "raeubertochter") |> 
  st_intersection(extent_raeubertochter)


p2 <- background + 
tm_shape(subset_raeubertochter_seg, bbox = extent_raeubertochter) +
  tm_symbols(size = 0.9, col = "segment_id", palette = my_palette2, alpha = 0.9, border.col = "NA",legend.col.show = FALSE) +
tm_shape(first_last_points_raeubertochter, bbox = extent_raeubertochter) +
  tm_symbols(size = 2, col = "gold", alpha = 1, border.col = "grey25", border.alpha = 0.5, legend.col.show = F) +
  tm_layout(frame = FALSE, bg.color = "transparent")



## save to png 
tmap_save(p2, "figures/start_endpoints_context.png", width = 250, height = 200, dpi = 300, units = "mm")


## Then show our buffer at the tunnel entry and exit
p3 <- background +  
tm_shape(buffer_entry, bbox = extent_raeubertochter) +
  tm_polygons(col = "red", border.col = "red", lwd = 0.5, alpha = 0.7) +
tm_shape(buffer_exit) +
  tm_polygons(col = "red", border.col = "red", lwd = 0.5, alpha = 0.7)+
tm_layout(frame = FALSE, bg.color = "transparent") 


## save to png 
tmap_save(p3, "figures/buffer.png", width = 250, height = 200, dpi = 300, units = "mm")


## Then show overlay -> whicht starting and endpoints lay within the entry/exit buffers
p4 <- background + 
tm_shape(buffer_entry, bbox = extent_raeubertochter) +
  tm_polygons(col = "red", border.col = "red", lwd = 0.5, alpha = 0.7) +
tm_shape(buffer_exit) +
  tm_polygons(col = "red", border.col = "red", lwd = 0.5, alpha = 0.7)+
  tm_shape(first_last_points_raeubertochter, bbox = extent_raeubertochter) +
   tm_symbols(size = 2, col = "gold", alpha = 1, border.col = "grey25", border.alpha = 0.5, legend.col.show = F) +
  tm_layout(frame = FALSE, bg.color = "transparent") 



## save to png 
tmap_save(p4, "figures/overlay.png", width = 250, height = 200, dpi = 300, units = "mm")


## then show unmerged/merged results
## unmerged
# p5 <- background +
#   tm_shape(subset_raeubertochter_seg, bbox = extent_raeubertochter) +
#   tm_symbols(size = 0.6, col = "segment_id", palette = my_palette2, alpha = 0.7, border.col = "NA", legend.col.show = FALSE) +
#   tm_layout(frame = FALSE, bg.color = "transparent") 
# 
# p5
# 
# ## save to png 
# tmap_save(p5, "raeubertochter_cor.png", width = 250, height = 200, dpi = 250, units = "mm")


## Adapt color palette to less segments
my_palette3 <- c("#9E0142", "#F46D43", "#FFFF66", "#3288BD",  "#5E4FA2")


## merged

load("rda_files/all_routes_seg_tunnel_cor.rda")
load("rda_files/raeubertochter_cor.rda")
load("rda_files/donner_cor.rda")

subset_raeubertochter_cor <- raeubertochter_cor[seq(1, nrow(raeubertochter_cor), by = 2), ]
subset_raeubertochter_cor <- st_intersection(subset_raeubertochter_cor, extent_raeubertochter)

subset_raeubertochter_cor$segment_id_cor <- as.numeric(subset_raeubertochter_cor$segment_id_cor)
subset_raeubertochter_cor$segment_id_cor <- as.factor(subset_raeubertochter_cor$segment_id_cor)


p6 <- background + 
tm_shape(subset_raeubertochter_cor, bbox = extent_raeubertochter) +
  tm_symbols(size = 0.9, col = "segment_id_cor", palette = my_palette3, alpha = 0.9, border.col = "NA", legend.col.show = FALSE) +
  tm_layout(frame = FALSE, bg.color = "transparent")

## show merged segments
print(p6)

## save to png 
tmap_save(p6, "figures/merged.png", width = 250, height = 200, dpi = 300, units = "mm")


```

## 5. Visualize Original and Alternative Trajectories

```{r visualize alternative routing,  fig.height= 9, fig.width=9, fig.align='center'}


load("rda_files/donner_cor.rda")
load("rda_files/raeubertochter_cor.rda")



## Create a new extent for this
coords <- rbind(c(2679514  , 1249013 ), c(2683075, 1249013 ), c(2683075, 1247635), c(2679514 , 1247635), c(2679514 , 1249013 ))


## Create a polygon
polygon <- st_polygon(list(coords))


## Convert to spatial object
extent2 <- st_sfc(polygon, crs = 2056)


## Clip spatial data to new extent to speed up computing time
streets_alt <- st_intersection(streets, extent2)
housing_alt <- st_intersection(housing, extent2)


## filter one representative delivery
donner_vis <- donner_cor |> 
   filter(segment_id_cor == 1)


## filter one representative alternative
alt <- alternative |> 
  filter(segment_id == 1) |> 
  group_by(segment_id, rank) |> 
  summarise(geometry = st_combine(geometry)) |> 
  st_cast("LINESTRING") |> 
mutate(
  rank = as.numeric(rank),
  rank = as.factor(rank)
)





## filter one representative start- and endpoint
df_odd_vis <- df_odd |> 
  filter(segment_id_cor == 1)

df_even_vis <- df_even |> 
  filter(segment_id_cor == 1)


## set mode to plot
tmap_mode("plot")


## create a new background
background2 <- tm_shape(streets_alt, bbox = extent2) +
 tm_lines(lwd = "width", scale = 2.5, legend.lwd.show = FALSE) +  
  tm_shape(housing_alt, bbox = extent2) +
  tm_polygons(col = "grey25", border.col = "white", lwd = 0.5, alpha = .9) 


## map original route
original_route <- background2 +
  tm_shape(donner_vis) +
  tm_dots(col = "#9E1FA2", size = .3, alpha = 0.8) +
 tm_layout(frame = FALSE, bg.color = "transparent") 



## print original routing
#print(original_route)

tmap_save(original_route, "figures/original_route.png", width = 300, height = 125, dpi = 300, units = "mm")


## combine original route with start and endpoints
tm4 <- background2 +
  tm_shape(donner_vis) +
  tm_dots(col = "#9E1FA2", size = .3, alpha = 0.8) +
  tm_shape(df_even_vis) +
  tm_dots(col = "steelblue1", size = 2) +
 tm_shape(df_odd_vis) +
  tm_dots(col = "gold", size = 2) +
 tm_layout(frame = FALSE, bg.color = "transparent")


## print map with start and endpoints
#print(tm4)


## save as png
 tmap_save(tm4, "figures/original_route_start_endpoints.png", width = 300, height = 125, dpi = 300, units = "mm")


my_palette4 <- c("#d7191c", "#fdae61","#2b83ba", "#acdfa9", "yellow")
 
 
 
## show alternative routing generated by hereR and based on starting and endpoinnts
tm5 <- background2 +
 tm_shape(alt) +
tm_lines(col = "rank", palette = my_palette4, lwd = 6, legend.col.show = FALSE, alpha = .8) +
 tm_shape(df_even_vis) +
  tm_dots(col = "steelblue1", size = 2) +
 tm_shape(df_odd_vis) +
  tm_dots(col = "gold", size = 2) +
 tm_layout(frame = FALSE, bg.color = "transparent") 


## show map
#print(tm5)


## save as png
tmap_save(tm5, "figures/alternative_routing.png", width = 300, height = 125, dpi = 300, units = "mm")


## combine everything together
tm6 <- background2 +
  tm_shape(alt) +
tm_lines( col = "rank", palette = my_palette4, lwd = 6, legend.col.show = FALSE, alpha = 0.8) +
   tm_shape(donner_vis) +
  tm_dots(col = "#9E1FA2", size = .25, alpha = 0.8) +
 tm_shape(df_even_vis) +
  tm_dots(col = "steelblue1", size = 2) +
 tm_shape(df_odd_vis) +
  tm_dots(col = "gold", size = 2) +
 tm_layout(frame = FALSE, bg.color = "transparent") 


## show combined map
#print(tm6)


## save as png
tmap_save(tm6, "figures/final_result.png", width = 300, height = 125, dpi = 300, units = "mm")



```

## 6. Visualize Map Matching

Interactive visualization of map matching results

```{r visualize map matching via tmap and interactive}

# Noch gedanken bez. dynamischeren visualisierungen machen

load("rda_files/filtered_df.rda")
load("rda_files/nearest.rda")

box = st_bbox(filtered_df)

tmap_mode("view")
tm_shape(street_network) + 
  tm_lines() +
tm_shape(nearest) + 
  tm_lines(col = "darkred", lwd = 2, lty = "dashed") +
tm_shape(filtered_df) + 
  tm_dots(col = "type") +
  tm_basemap(server = NULL) +
    tm_view(bbox =  box) 

```

static visualization via tmap and in a smaller spatial context

```{r visualize map matching via tmap and}

### Next, we want to visualize a subsample of our segmented data in a smaller spatial context
## Visualize Subsample in spatial context


## Define the coordinates of the polygon for a new extent
coords <- rbind(c(2682094, 1247370), c(2682094, 1247612), c(2682433, 1247612), c(2682433, 1247370), c(2682094, 1247370))


## Create a polygon
polygon <- st_polygon(list(coords))

## Convert to spatial object
extent4 <- st_sfc(polygon, crs = 2056)


## clip spatial data to new extent
housing1 <- st_intersection(housing, extent4)
streets1 <- st_intersection(streets, extent4)
surface1 <- st_intersection(surface, extent4)
combined_vis <- st_intersection(filtered_df, extent4)


## set mode
tmap_mode("plot")


## visualize map matching in a smaller spatial context
tm_7 <-  tm_shape(streets1, bbox = extent4) + 
   tm_lines(lwd = "width", scale = 2.5, legend.lwd.show = FALSE) +
  tm_shape(housing1, bbox = extent4) +
  tm_polygons(col = "grey25", border.col = "white", lwd = 0.5, alpha = 0.9) +
tm_shape(nearest) + 
  tm_lines(lwd = 1.5, lty = "solid", col = "grey") +
tm_shape(combined_vis$geometry) + 
  tm_dots(col = "purple", border.col = "purple", size = 0.4, alpha = .9, ) + 
tm_shape(combined) + 
  tm_dots(col = "orange", size = 0.3, alpha = .75, border.col = "orange",) +
  tm_layout(frame = FALSE, bg.color = "transparent") 


## Show map
print(tm7)


## save as png
 tmap_save(tm_7, "figures/map_matching.png", width = 200, height = 150, dpi = 1000, units = "mm")


```

## 7. Visualize uneven distribution

```{r}



## We want to visualize the problem here
## Problematic: uneven distribution alternative and original points as shown here
## first we filter delivery number 57 because it illustrates the problem very good
df_lines1 <- filtered_df |> 
  group_by(type, messenger, segment_id, rank) |> 
  summarise(geometry = st_combine(geometry)) |> 
  st_cast("LINESTRING")


## filter alternative segment 57
df_57_lines <- df_lines1 |> 
  filter(segment_id == 57 & (rank == 6 | rank == 7))


## filter original segment 57
df_57_points <- filtered_df |> 
  filter(segment_id == 57 & rank == 0)


## Define the coordinates of the new extent
coords <- rbind(c(2683114 , 1247977), c(2683114 , 1249227), c(2683785, 1249227), c(2683785, 1247977), c(2683114 , 1247977))

## Create a polygon
polygon <- st_polygon(list(coords))

## Convert to spatial object
extent <- st_sfc(polygon, crs = 2056)

## clip everything to new extent for faster computing
housing5 <- st_intersection(housing, extent)
streets5 <- st_intersection(streets, extent)
surface5 <- st_intersection(surface, extent)

combined_vis <- df_57_points |> 
  st_intersection(extent)

df_lines_vis <- df_57_lines |> 
  st_intersection(extent) 




## create a subset so only every second point is shown
subset_data <- combined_vis[seq(1, nrow(combined_vis), by = 2), ]
subset_data2 <- st_cast(df_lines_vis, "POINT")
subset_data2 <- subset_data2[seq(1, nrow(subset_data2), by = 2), ]

## set plotting mode
tmap_mode("plot")



problem <- tm_shape(streets5) +
  tm_lines(lwd = "width", scale = 2.5, legend.lwd.show = FALSE) +
tm_shape(housing5) +
  tm_polygons(col = "grey25", border.col = "white", lwd = 0.5, alpha = .9) +
tm_shape(subset_data) +
  tm_dots(size = 0.8, col = "#9E1FA2", alpha = 0.8, border.col = "grey25", border.alpha = 0.2) +
  tm_shape(subset_data2) +
  tm_symbols(size = 0.8, col = "rank", palette = c("#11E2A1", "#11E2A1"), alpha = 0.8, border.col = "grey25", border.alpha = 0.2, legend.col.show = FALSE) +
  tm_layout(frame = FALSE) 
 

## show map
print(problem)


## save as png
tmap_save(problem, "figures/problem.png", width = 125, height = 225, dpi = 300, units = "mm")


load("rda_files/overview.rda")

overview$type <- factor(overview$type, levels = c("original", "alternative"))

uneven <- ggplot(overview) +
  geom_boxplot(aes(x = type, y = count, fill = type), alpha = 0.75) +
  scale_fill_manual(values = c("#9E1FA2", "#11E2A1")) +
  labs(x = "") +
  theme_minimal() +
  theme(
    legend.position = "none"
  )


ggsave(plot = uneven, "figures/uneven.png", width = 300, height = 150, dpi = 500, units = "mm")

```

# 8. Visualize Raw vs. Interpolated

```{r comparison of raw and interpolated fixes}



## We want to compare the raw gps and interpolaed data
## Define the coordinates of the polygon
coords <- rbind(c(2683114 , 1247977), c(2683114 , 1249227), c(2683785, 1249227), c(2683785, 1247977), c(2683114 , 1247977))

## Create a polygon
polygon <- st_polygon(list(coords))


## Convert to spatial object
extent <- st_sfc(polygon, crs = 2056)


## Clip to new extent
housing5 <- st_intersection(housing, extent)
streets5 <- st_intersection(streets, extent)
surface5 <- st_intersection(surface, extent)



## select interpolated data
df_57_interp <- df_single_points |> 
  filter(segment_id == 57 & (rank == 0 | rank == 6 | rank == 7))

df_57_interp$type <- factor(df_57_interp$type, levels = c("original", "alternative"))
df_57_interp$rank <- as.factor(df_57_interp$rank)


## clip to new extent
df_57_interp <- df_57_interp |> 
  st_intersection(extent)


## create subset so only every second point is shown
subset_data_interp <- df_57_interp[seq(1, nrow(df_57_interp), by = 2), ]


my_palette6 = c( "#9E1FA2","#11E2A1","#11E2A1")

tmap_mode("plot")

problem_solved <- tm_shape(streets5) +
  tm_lines(lwd = "width", scale = 2.5, legend.lwd.show = FALSE) +
tm_shape(housing5) +
  tm_polygons(col = "grey25", border.col = "white", lwd = 0.5, alpha = .9) +
tm_shape(subset_data_interp) +
  tm_symbols(size = 0.8, col = "rank", palette = my_palette6, alpha = 0.8, border.col = "grey25", border.alpha = 0.2, legend.col.show = FALSE) +
  tm_layout(frame = FALSE) 
 

## show map
print(problem_solved)


## save as png
tmap_save(problem_solved, "figures/problem_solved.png", width = 125, height = 225, dpi = 300, units = "mm")




```

## 9. Visualize alternatives and originals

```{r visualize original and alternatives per messenger, fig.width=12, fig.height=12}

## simple overview alternatives vs originals
df_single_points$type <- factor(df_single_points$type, levels = c("original", "alternative"))


## split dataframe into different messengers
raeubertochter_final <- filter(df_single_points, messenger == "raeubertochter")
donner_final <- filter(df_single_points, messenger == "donner")


## create own color palette
my_palette <- colorRampPalette(brewer.pal(11, "Spectral"))


## raeubertochter
raeubertochter_final <- ggplot() +
  geom_sf(data = outline, alpha = 0) +
  geom_sf(data = raeubertochter_final, aes(color = type), size = 0.6) +
   scale_color_manual(values = c( "#9E1FA2", "#11E2A1")) +
  coord_sf(datum = st_crs(2056)) +
  facet_grid(~type) +
  labs(x = "E",
       y = "N") +
    theme_minimal() +
  theme(
    legend.position = "none"
  ) +
  labs(title = "Raeubertochter")

ggsave(plot = raeubertochter_final, "figures/raeubertochter_final.png", width = 100, height = 100, dpi = 300, units = "mm")




## Donner
donner_final <- ggplot() +
  geom_sf(data = outline, alpha = 0) +
  geom_sf(data = donner_final, aes(color = type), size = 0.6) +
   scale_color_manual(values = c( "#9E1FA2", "#11E2A1")) +
  coord_sf(datum = st_crs(2056)) +
  facet_grid(~type) +
  labs(x = "E",
       y = "N") +
    theme_minimal() +
  theme(
    legend.position = "none"
  ) +
  labs(title = "Donner")


ggsave(plot = donner_final, "figures/donner_final.png", width = 100, height = 100, dpi = 300, units = "mm")


```

# 10. Visualize Segment length

```{r}

load("rda_files/new.rda")

length_comp <- new |> 
  select(type, messenger, segment_id, rank,  tot_segment_length) |>


df_spread <- length_comp |> 
  group_by(type, messenger, segment_id, rank) |> 
  summarise(
    tot_travel = mean(tot_segment_length)
  ) |> 
  mutate(
    rank = factor(rank, levels = c(0,1,2,3,4,5,6,7))
  )

## How differ segment length between messengers? only original routes considered

df_spread_original <- df_spread |> 
  filter(type == "original")


seg_length <- ggplot(df_spread_original, aes(x = messenger, y = tot_travel, fill = messenger)) +
  geom_boxplot(alpha = 0.75) +
   labs(y = "total segment length [m]") +
  scale_fill_manual(values = c("gold", "forestgreen")) +
   theme_classic() +
  theme(
    legend.position = "none", 
     axis.title.y = element_text(margin = margin(r = 20)),
    axis.title.x = element_blank()
  ) +
  labs(title = "")

ggsave(plot = seg_length, "figures/seg_length.png", width = 300, height = 150, dpi = 500, units = "mm")



## Are there differences between shortest route generated and route choosen?
comparison1 <- df_spread |> 
  st_drop_geometry() |> 
  filter(rank == 0 | rank == 1) |> 
  group_by(segment_id, rank) |> 
  summarise(mean = mean(tot_travel)) |> 
  spread(key = rank, value = mean) |> 
  rename("original" = "0" , "alternative1" = "1")

comp <- pivot_longer(comparison1, -segment_id)


shortest <- ggplot(comp, aes(x = name, y = value, fill = name,   alpha = 0.75)) +
  geom_boxplot() +
   labs(y = "total segment length [m]") +
  scale_fill_manual(values = c("#11E2A1","#9E1FA2")) +
   theme_classic() +
  theme(
    legend.position = "none",
    axis.title.y = element_text(margin = margin(r = 20)),
    axis.title.x = element_blank())
  
ggsave(plot = shortest, "figures/shortest.png", width = 300, height = 150, dpi = 500, units = "mm")

## How do segment lengths differ in contrast to alternatives? 

seg_length_mess <- ggplot(df_spread, aes(x = messenger, y = tot_travel, fill = type,   alpha = 0.75)) +
  geom_boxplot() +
   labs(y = "total segment length [m]") +
  scale_fill_manual(values = c("#11E2A1","#9E1FA2")) +
   theme_classic() +
  theme(
    legend.position = "none",
    axis.title.y = element_text(margin = margin(r = 20)),
    axis.title.x = element_blank()
  ) +
  labs(title = "")

ggsave(plot = seg_length_mess, "figures/seg_length_mess.png", width = 300, height = 150, dpi = 500, units = "mm")



# We see that they differ slightly, raeubertochter more than donner.. but how?


df_spread$rank1 <- factor(df_spread$rank, labels =  c("Original"  ,"A1", "A2", "A3", "A4", "A5", "A6", "A7"), levels = c( "0", "1", "2", "3", "4", "5", "6", "7"))


anova <- ggplot(df_spread, aes(x = rank1, y = tot_travel, fill = type)) +
 geom_jitter(aes(colour = tot_travel), alpha = 0.25) +
  geom_boxplot(alpha = 0.75) +
   labs( y = "total segment length [m]") +
  scale_fill_manual(values = c("#11E2A1","#9E1FA2")) +
    scale_color_gradient(low = "gray", high = "black") +  # Color gradient for jittered points
   theme_classic() +
  theme(
    legend.position = "none",
        axis.title.y = element_text(margin = margin(r = 20)),
    axis.title.x = element_blank()
  ) 

ggsave(plot = anova, "figures/anova.png", width = 300, height = 150, dpi = 500, units = "mm")

## visualize anova
rank <- ggplot(df_spread, aes(x = rank1, y = tot_travel, fill = type,  alpha = 0.75)) +
  geom_boxplot() +
   labs(y = "total segment length [m]") +
  scale_fill_manual(values = c("#11E2A1","#9E1FA2")) +
   theme_classic() +
  theme(
    legend.position = "none"
  ) +
  labs(title = "")

```

# 11. Visualize street width

```{r}


width_preference <- new |>
  st_drop_geometry() |> 
  group_by(type, segment_id, rank, width) |> 
  summarise(
    sum = sum(travel_dist, na.rm = T)
  ) 

##Create a complete set of segment IDs and street widths, credits go to: https://stackoverflow.com/questions/73717164/create-all-combinations-of-two-variables-from-two-dataframes-while-keeping-all-o
complete_set <- expand.grid(segment_id = unique(width_preference$segment_id),
                            rank = unique(width_preference$rank),
                           width = as.factor(c(1,2,3,4,6,8)),
                           type = as.factor(c("original", "alternative")))
                           #messenger = as.factor(c("raeubertochter", "donner")))

complete_set <- complete_set |> 
  filter(!(type == "original" & rank != 0)) |> 
  filter(!(type == "alternative" & rank == 0)) 

##Left join this set with your original data frame
complete_df <-  full_join(complete_set, width_preference, by = c("type","segment_id", "rank", "width"))


#Replace NA values with 0
complete_df[is.na(complete_df)] <- 0


width_preference2 <- complete_df |> 
   group_by(type, segment_id, rank) |>
  mutate(
    width = width,
    sum = as.numeric(sum),
    n = n(),
    tot = sum(sum),
    perc = (sum / tot) * 100
  ) |> 
  na.omit()


width_preference2 <- width_preference2 |> 
   group_by(type, segment_id, rank) |>
  mutate(
    sum_tot = sum(perc)
  )


choice_avg <- width_preference2 |> 
   group_by(type, segment_id, width) |>
 summarise(
   mean_perc = mean(perc)
  ) 

all_together <- choice_avg |> 
  group_by(type, width) |> 
   summarise(
   mean_perc = mean(mean_perc)
  ) 


spread <- spread(all_together, key = type, value = mean_perc)
spread$diff <- spread$original - spread$alternative


spread <- spread[order(spread$diff),]
spread$width <- factor(spread$width, levels = spread$width,
                       labels = c("1m street", "6m street", "8m street",
                                  "4m street", "2m street", "3m street"))

preference <- ggplot(spread, aes(x = width, y = diff)) +
  geom_segment(aes(x = width, xend = width, y = 0, yend = diff), color = "grey60", lwd = 1.5) +
  geom_point(size = 5, pch = 21, bg = "#9E1FA2", color = "#9E1FA2") +
  coord_flip() +
  theme_minimal() +
  labs(
    y = "← Less Preferred | Preferred →                ",
    x = ""
  ) +
  theme(
    axis.text.x = element_blank()
  )

ggsave(plot = preference, "figures/preference.png", width = 300, height = 150, dpi = 500, units = "mm")

```

# 12. Visualize Slope

```{r}


# max averae slope per 100m
# mean average_slope per 100m


# counts per category / m per categor

slope <- new |> 
  group_by(type, segment_id, rank, index) |> 
  na.omit() |> 
  summarise(
    n = n(),
     mean_slope_up = mean(ifelse(slope_calc > 0, slope_calc, 0), na.rm = TRUE),
    max_slope_up = max(ifelse(slope_calc > 0, slope_calc, 0), na.rm = TRUE),
      mean_slope = mean((slope_calc), na.rm = TRUE),
     max_slope = max((slope_calc), na.rm = TRUE))



# density plot of the results
results1 <- results |> 
  group_by(type, segment_id) |> 
  filter(mean_slope100 == max(slope_up))



slope$rank1 <- factor(slope$rank, labels =  c("Original"  ,"A1", "A2", "A3", "A4", "A5", "A6", "A7"), levels = c( "0", "1", "2", "3", "4", "5", "6", "7"))



slope_vis <- slope |> 
  filter(mean_slope_up < 10)

slope_vis2 <- slope |> 
  filter(max_slope_up < 30)





mean_slope_up <-  ggplot(slope_vis, aes(x = mean_slope_up, y = rank1, height = ..density..)) +
  geom_density_ridges(aes(fill = rank), alpha = 0.75, stat = "density") +
  scale_fill_manual(values =c("#11E2A1","#11E2A1","#11E2A1","#11E2A1","#11E2A1","#11E2A1","#11E2A1","#9E1FA2")) +
 theme_minimal() +
    labs(
    x = "mean positive slope per 100m subsegments [%]"
  ) +
  theme(
    legend.position = "none",
    axis.title.x = element_text(margin = margin(t = 20)),
    axis.title.y = element_blank()
    
  )

ggsave(plot = mean_slope_up, "figures/mean_slope_up.png", width = 300, height = 150, dpi = 500, units = "mm")


max_slope_up <- ggplot(slope_vis2, aes(x = max_slope_up, y = rank1, height = ..density..)) +
  geom_density_ridges(aes(fill = rank), alpha = 0.75, stat = "density") +
scale_fill_manual(values =c("#11E2A1","#11E2A1","#11E2A1","#11E2A1","#11E2A1","#11E2A1","#11E2A1","#9E1FA2")) +
 theme_minimal() +
   labs(
    x = "max positive slope per 100m subsegments [%]") +
  theme(
    legend.position = "none",
    axis.title.x = element_text(margin = margin(t = 20)),
    axis.title.y = element_blank()
  )


ggsave(plot = max_slope_up, "figures/max_slope_up.png", width = 300, height = 150, dpi = 500, units = "mm")



# create counts per category

slope_long <- slope |> 
  select(type, segment_id, rank, index, mean_slope_up)


slope_result <- slope |> 
  st_drop_geometry() |> 
  group_by(type, segment_id, rank) |> 
  summarise(
    kat1 = sum(mean_slope_up > 2 & mean_slope_up <= 4),
    kat2 = sum(mean_slope_up > 4 & mean_slope_up <= 6),
    kat3 = sum(mean_slope_up > 6)
      )




slope_sum <- pivot_longer(slope_result, -c(type, segment_id, rank))

slope_sum$name <- factor(slope_sum$name, levels = c("kat1", "kat2", "kat3"), labels = c("2-4%", "4-6%", ">6%"))

slope_final <- slope_sum |> 
  group_by(type, segment_id, name) |> 
  summarise(
      mean_count = mean(value) )




slope_category <- ggplot(slope_final, aes(x = name, y = mean_count, fill = type), alpha = 0.75) +
  geom_boxplot() +
  scale_y_log10() +
   labs(y = "mean count of subsegments per delivery (Log Scale)",
        x = "slope categories") +
  scale_fill_manual(values = c("#11E2A1","#9E1FA2")) +
   theme_classic() +
  theme(
       axis.title.y = element_text(margin = margin(r = 20)),
    axis.title.x = element_text(margin = margin(t = 20))
  )

ggsave(plot = slope_category, "figures/category.png", width = 300, height = 150, dpi = 500, units = "mm")



```
